---
title: "Initial Korean covid 19"
author: "Ixcell Espinoza- Alvarado"
date: "Spring 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

# Introduction

$h_0$ There is no difference in death rate based on sex

$h_0$ There is a difference in death rates based on the sex of the person 


# Methods

Variables for this study are sex and result

We will find if there is a difference in death rates based on the sex 

Results is categorical variable 

Sex is a categorical variable so we will use the cat~cat helper file 

## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```


  The graph shows that the is a difference in death rates based on sex. Looking at the graph above we can see that Males have a higher death rate then Females.with the average for males being 60% and females average being below 40% 




## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The table shows that males death rates are higher then the females males having a rate of 64% while females having a rate of 36%


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi-square test p value is 0.0012 with is less than 0.05 therefore we will reject the hypothesis meaning there is a significant difference in deaths based on sex
the p-value on the fisher exact is 0.0009 meaning the null hypothesis can be rejected because it is less then 0.05
the odds ratio is 2.3:1 this means men are at higher risk then women due to their larger death rate 

## Conclusions

"Is there a difference in korean data death rate based on sex?"
with the numerical graphical results we can come to the conclusion that the sex of the person does matter as men do have a higher mortality rate. Also because our P value = 0.0012 is smaller than 0.05 the null hypothesis is rejected and proves that there is a difference in death rated due to the persons sex. 
